---
title: "Multivariate_Analysis"
author: "Ishana Rana"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Multivariate Analysis
The goal of this part is to analyze potential inter-metric correlations.

## Loading datasets:
```{r}
bertscore <- read.csv("C:\\Users\\babus\\OneDrive\\Documents\\uni uzh\\FS25\\conversational speech processing\\mypaper\\Beyond-WER-in-ASR\\data\\eval_results\\BERTScore_scores.csv", skip=1)
str(bertscore)
head(bertscore)
summary(bertscore)
```

```{r}
bleu <- read.csv("C:\\Users\\babus\\OneDrive\\Documents\\uni uzh\\FS25\\conversational speech processing\\mypaper\\Beyond-WER-in-ASR\\data\\eval_results\\BLEU_scores.csv")
str(bleu)
head(bleu)
summary(bleu)
```

```{r}
# splitting dataset as it contains scores of ROUGE-1 and ROUGE-L
lines <- readLines("C:\\Users\\babus\\OneDrive\\Documents\\uni uzh\\FS25\\conversational speech processing\\mypaper\\Beyond-WER-in-ASR\\data\\eval_results\\ROUGE_scores.csv")
split_index <- grep("ROUGE-L", lines)

rouge1_lines <- lines[2:(split_index - 1)]
rougel_lines <- lines[(split_index + 1):length(lines)]

rouge1 <- read.csv(text = rouge1_lines)
rougel <- read.csv(text = rougel_lines)

str(rouge1)
head(rouge1)
summary(rouge1)
tail(rouge1_lines, 1)
rouge1 <- rouge1[-nrow(rouge1), ]


str(rougel)
head(rougel)
summary(rougel)
```

```{r}
wer <- read.csv("C:\\Users\\babus\\OneDrive\\Documents\\uni uzh\\FS25\\conversational speech processing\\mypaper\\Beyond-WER-in-ASR\\data\\eval_results\\WER_scores.csv", skip=1)
str(wer)
head(wer)
summary(wer)
```

## Analysis:
```{r}
library(ggplot2)
#install.packages("corrplot")
library(corrplot)
```

```{r}
# specifying directory where the created plots shall be saved at
plot_dir = "C:\\Users\\babus\\OneDrive\\Documents\\uni uzh\\FS25\\conversational speech processing\\mypaper\\Beyond-WER-in-ASR\\data\\eval_results\\R_code\\for_paper"
dir.create(plot_dir, showWarnings = FALSE)
```

### Correlation Matrix

#### Sidenote
Because of the small sample size these results also lack statistical power.

#### with punctuation:
```{r}
metrics_df_with_punct <- data.frame(
  File = wer$File,
  WER = wer$with_punct,
  BLEU = bleu$with_punct,
  ROUGE1 = rouge1$with_punct,
  ROUGEL = rougel$with_punct,
  BERTScore = bertscore$with_punct
)
metrics_df_with_punct
```

```{r}
# Compute correlation matrix (Spearman is more robust for small samples)
cor_matrix_with_punct <- cor(metrics_df_with_punct[,-1], method = "spearman")
print(cor_matrix_with_punct)
# save output to image:
png(file.path(plot_dir, "with_punct_corr_matrix_circle.png"), width = 1000, height = 800, res = 150)
corrplot(cor_matrix_with_punct, method = "circle", type = "upper", tl.cex = 0.8)
dev.off()

png(file.path(plot_dir, "with_punct_corr_matrix_number.png"), width = 1000, height = 800, res = 150)
corrplot(cor_matrix_with_punct, method = "number", type = "upper", tl.cex = 0.8)
dev.off()

```

#### without punctuation:
```{r}
metrics_df_without_punct<- data.frame(
  File = wer$File,
  WER = wer$without_punct,
  BLEU = bleu$without_punct,
  ROUGE1 = rouge1$without_punct,
  ROUGEL = rougel$without_punct,
  BERTScore = bertscore$without_punct
)
metrics_df_without_punct
```

```{r}
cor_matrix_without_punct <- cor(metrics_df_without_punct[,-1], method = "spearman")
print(cor_matrix_without_punct)

# save output to image:
png(file.path(plot_dir, "without_punct_corr_matrix_circle.png"), width = 1000, height = 800, res = 150)
corrplot(cor_matrix_without_punct, method = "circle", type = "upper", tl.cex = 0.8)
dev.off()

png(file.path(plot_dir, "without_punct_corr_matrix_number.png"), width = 1000, height = 800, res = 150)
corrplot(cor_matrix_without_punct, method = "number", type = "upper", tl.cex = 0.8)
dev.off()
```


# Testing Correlation Significance
#### with punctuation:

```{r}
cor.test(metrics_df_with_punct$WER, metrics_df_with_punct$BERTScore, method = "spearman")
cor.test(metrics_df_with_punct$WER, metrics_df_with_punct$ROUGE1, method = "spearman")
cor.test(metrics_df_with_punct$WER, metrics_df_with_punct$ROUGEL, method = "spearman")
cor.test(metrics_df_with_punct$WER, metrics_df_with_punct$BLEU, method = "spearman")

```
The results show that WER has:
- strong negative relationship with BERTScore but is statistically not significant
- has moderate negative correlations with ROUGE-1, ROUGE-L and BLEU but is statistically also not significant

```{r}
cor.test(metrics_df_with_punct$BLEU, metrics_df_with_punct$ROUGE1, method = "spearman")
cor.test(metrics_df_with_punct$BLEU, metrics_df_with_punct$ROUGEL, method = "spearman")
```
These results show that BLEU has:
- perfect positive correlation with ROUGE-1 that is also statistically significant
- moderate strong positive relation with ROUGE-L but not statistically significant

```{r}
cor.test(metrics_df_with_punct$ROUGEL, metrics_df_with_punct$BERTScore, method = "spearman")
```
This test shows that there is no correlation at all, indicating they capture different linguistic phenomena.



#### Without punctuation:
For this condition, the pairs that showed some difference to the "with_punct" ones were tested. None of them showed any statistically significant correlation.
```{r}
cor.test(metrics_df_without_punct$WER, metrics_df_without_punct$BLEU, method = "spearman")
```

```{r}
cor.test(metrics_df_without_punct$BLEU, metrics_df_without_punct$ROUGEL, method = "spearman")
cor.test(metrics_df_without_punct$BLEU, metrics_df_without_punct$BERTScore, method = "spearman")
```

```{r}
cor.test(metrics_df_without_punct$ROUGEL, metrics_df_without_punct$BERTScore, method = "spearman")
```


# Plotting Images for paper:
```{r}
metrics_df_with_punct <- data.frame(
  File = wer$File,
  WER = wer$with_punct,
  BLEU = bleu$with_punct,
  ROUGE1 = rouge1$with_punct,
  ROUGEL = rougel$with_punct,
  BERTScore = bertscore$with_punct
)
metrics_df_with_punct

metrics_df_without_punct<- data.frame(
  File = wer$File,
  WER = wer$without_punct,
  BLEU = bleu$without_punct,
  ROUGE1 = rouge1$without_punct,
  ROUGEL = rougel$without_punct,
  BERTScore = bertscore$without_punct
)
metrics_df_without_punct


cor_matrix_with_punct <- cor(metrics_df_with_punct[,-1], method = "spearman")
cor_matrix_without_punct <- cor(metrics_df_without_punct[,-1], method = "spearman")

png(file.path(plot_dir, "combined_corr_matrix_circle.png"), width = 1000, height = 1600, res = 150)
par(mfrow = c(2, 1))  # 2 rows, 1 column layout

corrplot(cor_matrix_with_punct, method = "circle", type = "upper", tl.cex = 0.8,
         title = "With Punctuation", mar = c(0,0,2,0))  # add margin space for title

corrplot(cor_matrix_without_punct, method = "circle", type = "upper", tl.cex = 0.8,
         title = "Without Punctuation", mar = c(0,0,2,0))

dev.off()


```